---
title: "Untitled"
author: "Bilal Gilani"
date: "2/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## LOAD DATA
```{r}
load("GA_VOTERS.rda")
```

## RENAME/EDIT COLUMNS
```{r}
library(tidyverse)

colnames(GA_voter)[colnames(GA_voter)=="YEAR"]<-"REG_YEAR"

GA_voter <- GA_voter %>%
  mutate(AGE_2021 = 2021 - BIRTHDATE) ## Age in 2021?
```

## COUNTY INFO
```{r}
library(zipcodeR)

test <- search_state("GA")
test <- test[, c("zipcode", "county")] ## only want county

## Need to change the column name of "ZIP5" in order to acquire County name
colnames(GA_voter)[colnames(GA_voter)=="ZIP5"] <- "zipcode"

## convert from factor to character
as.character(GA_voter$zipcode)
GA_voter[["zipcode"]]<-as.character(GA_voter[["zipcode"]])

## combine
GA_voter <- left_join(GA_voter, test, by = "zipcode")

table(GA_voter$county) ## Two DeKalb counties

which(GA_voter$county == "Dekalb County")

GA_voter$county[GA_voter$county == "Dekalb County"] <- "DeKalb County" ## create ONE DeKalb County
which(GA_voter$county == "Dekalb County") ## FIXED

## Check if any counties are missing
sum(is.na(GA_voter$county))

## create new dataset without NA's
GA_voter_full <- GA_voter[!is.na(GA_voter$county),]
sum(is.na(GA_voter_full$county))

## change the county variable to mimic the tidycensus county variable later in the code
GA_voter_full$county <- paste(GA_voter_full[["county"]], "Georgia", sep = ", ")

## Registrations per year (ALL)
GA_year_all <- as.data.frame(table(GA_voter_full$REG_YEAR)) ## Make note of 1965-1966 and 1975-1976
colnames(GA_year_all)[colnames(GA_year_all)=="Freq"] <- "Reg_Voters"

GA_year_sub <- GA_year_all %>%          ## Use for Presentation
  slice(45:61)

## Create 2012-2016 Subset
GA_voter16 <- subset(GA_voter_full, GA_voter_full$REG_YEAR >= 2012 & GA_voter_full$REG_YEAR <= 2016) ## DO I NEED TO SUBSET? - YES

## Create County registrations
GA_county <- table(GA_voter16$county)
GA_county <- as.data.frame(GA_county)

colnames(GA_county)[colnames(GA_county)=="Var1"] <- "County"

colnames(GA_county)[colnames(GA_county)=="Freq"] <- "Reg_Voters"

GA_county <- GA_county[-c(32),] ## Remove Cleburn county

## Registrations per year (2012-2016)
GA_year16 <- as.data.frame(table(GA_voter16$REG_YEAR))

colnames(GA_year16)[colnames(GA_year16)=="Freq"] <- "Reg_Voters"

summary(GA_voter16$BIRTHDATE)
 
```

## TIDYCENSUS
```{r}
library(dplyr)
library(ggplot2)
library(knitr)
library(psych)
library(tidycensus)

census_api_key("79dcdbc724a9939f8ec75b6642f1e59f57da70e9")

v16 <- load_variables(2016, "acs5", cache = TRUE)

##  NOTE: Should create vector of desired variables
```

Desired variables:

Population by Race:
  Total: B02001_001
  White: B02001_002	
  Black: B02001_003 
  Native American: B02001_004
  Asian: B02001_005 
  Hawaiian/Pacific Islander: B02001_006
  Some **other** race alone: B02001_007
  Two or More Races: B02001_008

Median Income: B19013_001
Median Age: B01002_001

Citizenship:
  Total: B05001_001
  U.S. Citizen born in U.S.: B05001_002
  U.S. Citizen born in PR or Islands: B05001_003
  U.S. Citizen born abroad of American parent(s): B05001_004
  **Naturalized** U.S. Citizen: B05001_005

English Literacy:
  Total: B06007_001
  Speak Only English: B06007_002
  Speak Only Spanish: B06007_003
  Speak Spanish/Speak English "very well": B06007_004
  Speak Spanish/Speak English less than "very well": B06007_005
  Speak other languages: B06007_006 
  Speak other languages/Speak English "very well" : B06007_007
  Speak other languages/Speak English less than "very well" : B06007_008

Education:
  Total: B06009_001
  Less than high school graduate: B06009_002
  High school graduate (includes equivalency): B06009_003 
  Some college or Associate's degree: B06009_004 
  Bachelor's degree: B06009_005
  Graduate or professional degree: B06009_006
  


## COUNTY
```{r}
## TEST: Georgia 2016, NOTE: acs is taking data from a 5 year period of 2012-2016
test2 <- get_acs(geography = "county",
                  variables = "B02001_001",
                  state = "GA",
                   year = 2016)

## Create vector of desired variables
variable_vec <- c("B01003_001", ## Total Population
                  "B01002_001", ## Median Age
                  "B19013_001", ## Median income
                  paste0("B02001_00", 1:8), ## Race variables
                  paste0("B05001_00", 1:6), ## Citizenship variables
                  paste0("B06007_00", c(1:3, 6)), ## Language variables
                  paste0("B06009_00", 1:6) ## Education variables
                  )

## Use vector
GA2016 <- get_acs(geography = "county",
                  variables = variable_vec,
                  state = "GA",
                   year = 2016)

library(reshape)
GA2016 <- cast(GA2016, NAME~variable, value = "estimate")

## Rename columns
colnames(GA2016) <- c("County", "Median Age", "Gen. Population",
                      ## Race variables
                      "Race_Total", "White", "Black", "Native American", "Asian", "Hawaiian/Pacific Islander", "Other", "Two or More Races",
                      ## Citizenship variables
                      "Citizenship_Total", "U.S. Citizen born in U.S", "U.S. Citizen born in PR or Islands", 
                      "U.S. Citizen born abroad of American parent(s)","Naturalized U.S. Citizen", "Not a U.S. Citizen",
                      ## Language variables
                      "Language_Total", "Speak Only English", "Speak Spanish", "Speak other languages",
                      ## Education variables
                      "Education_Total", "Less than high school graduate", "High school graduate (includes equivalency)", 
                      "Some college or Associate's degree", "Bachelor degree", "Graduate or professional degree",
                      "Median Income"
                      )
                      

## Need to re-attach GEOID as well as add registered voters by county
test2 <- test2[, c("GEOID", "NAME")]
colnames(test2) <- c("GEOID", "County")

GA2016 <- left_join(GA2016, test2, by = "County")
GA2016 <- left_join(GA2016, GA_county, by = "County")

## Reorder columns
GA2016 <- GA2016[, c("GEOID", colnames(GA2016)[!colnames(GA2016) %in% c("GEOID")])] ## Move GEOID to first column

GA16_model <- GA2016[, colnames(GA2016)!="GEOID"] ## USE THIS FOR MODELING

```

```{r}
## Which counties are the most diverse?

diversity <- GA16_model[, c("County", "Race_Total", "White", "Black", "Native American", "Asian",
                            "Hawaiian/Pacific Islander", "Other", "Two or More Races")]

diversity <- diversity %>%
  mutate(Minority_Percent = (Black + `Native American` + Asian + `Hawaiian/Pacific Islander` + 
                               `Two or More Races` + diversity$Other) / `Race_Total`)

## Least diverse?
diversity <- diversity[order(diversity$Minority_Percent),]
head(diversity, n = 10)

## Most diverse?
diversity <- diversity[order(-diversity$Minority_Percent),]
head(diversity, n = 10)

## In both of the general/presidential election cases of 2012 and 2016, the Democratic candidate won in the top 10 most diverse counties. 
## Whereas,conversely, the Republican candidate won in the bottom 10 diverse counties.
```



USE GA16_model DATASET
## MODELS
```{r}
## Correlations?
GA16_cor <- select(GA16_model, -c(County))
cor_GA <- cor(GA16_cor)

library(Hmisc)
cor_GA2 <- rcorr(as.matrix(cor_GA))

cor_GA2

library(corrplot)
corrplot(cor_GA, type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45)



```

```{r}
## TEST:  Median Income model?
attach(GA16_model)
income.lm <- lm(`Median Income` ~. -County, data = GA16_model)

summary(income.lm)

income.lm

plot(income.lm)

```


```{r}
## Voter Registration

library(data.table)

## Drop county 
x <- GA16_model[, !(colnames(GA16_model) %in% c("County"))]
x <- setDT(x) # Convert x to a data table

## Create a vector of ratio variables by dividing each category by total observations per category

## Race variables
race_cols <- c("White", "Black", "Native American", "Asian", "Hawaiian/Pacific Islander",
               "Other", "Two or More Races")
x[, (race_cols) := lapply(.SD, function(x) {x / Race_Total}), .SDcols = race_cols] ## Divide each variable by Gen. Population to get ratios

## Education variables
education_cols <- c("Less than high school graduate", "High school graduate (includes equivalency)", 
                    "Some college or Associate's degree", "Bachelor degree", "Graduate or professional degree")
x[, (education_cols) := lapply(.SD, function(x) {x / Education_Total}), .SDcols = education_cols] ## Divide each variable by Gen. Population to get ratios

## Citizenship variables
citizenship_cols <- c("U.S. Citizen born in U.S", "U.S. Citizen born in PR or Islands", 
                      "U.S. Citizen born abroad of American parent(s)","Naturalized U.S. Citizen", "Not a U.S. Citizen")
x[, (citizenship_cols) := lapply(.SD, function(x) {x / Citizenship_Total}), .SDcols = citizenship_cols] ## Divide each variable by Gen. Population to get ratios

## Language variables
language_cols <- c("Speak Only English", "Speak Spanish", "Speak other languages")
x[, (language_cols) := lapply(.SD, function(x) {x / Language_Total}), .SDcols = language_cols] ## Divide each variable by Gen. Population to get ratios


## Calculate percentage of registered voters (dependent variable for model)
x[, "Reg_Percent" := Reg_Voters / `Gen. Population`]

## Drop Gen. Population, Race_Total, Citizenship_Total, Education_Total, and Language_Total
x <- x[, !colnames(x) %in% c("Gen. Population", "Race_Total", "Citizenship_Total", "Education_Total", "Language_Total"), with = FALSE]

## Have to drop one category from each categorical variable to avoid perfect multicollinearity
## The resulting model coefficients will be relative to the following baseline:
##     Two or more races
##     Not a US citizen
##     Speak other languages
##     Graduate or professional degree
x <- x[, !colnames(x) %in% c("Two or More Races", "Not a U.S. Citizen", "Speak other languages", "Graduate or professional degree"), with = FALSE]

library(lmtest)
library(sandwich)
library(car)

## Fit fractional logistic regression using glm function with binomial family
glm1 <- glm(Reg_Percent ~ . - Reg_Voters,  data = x, family = binomial)
plot(glm1)

## Calculate robust standard errors
se_glm_robust = coeftest(glm1, vcov = vcovHC(glm1, type="HC"))
print(se_glm_robust) 

## Test for collinearity
model_vif <- vif(glm1)
print(model_vif)
## There appears to be multicollinearity among the independent variables as indiciated by the variance inflation factors
## However, there is strong intuition for including these categorical variables in the model, and we have already
## accounted for perfect collinearity within categorical variables.

## Check if residual mean is statistically different from zero
t.test(glm1[["residuals"]])
## The mean of the residuals is not statistically different from zero


## Residual normality tests
## Plot Q-Q normality plot
qqnorm(glm1[["residuals"]], pch = 1, frame = FALSE) 
qqline(glm1[["residuals"]], col = "steelblue", lwd = 2)
## Q-Q plots indicate that the residuals are not normally distributed

## Shapiro-Wilk test
## Null hypothesis: Residuals are normally distributed
## Alternate hypothesis: Residuals are not normally distributed
## p-vale threshold: 0.05 
sw_test <- shapiro.test(glm1[["residuals"]])
print(sw_test)
## p-value is <0.05, so we reject the null hypothesis that the residuals are normally distributed

## Kolmogorov-Smirnov test
## Null hypothesis: Residuals are normally distributed
## Alternate hypothesis: Residuals are not normally distributed
## p-vale threshold: 0.05 
ks_test <- ks.test(glm1[["residuals"]], "pnorm", mean = mean(glm1[["residuals"]]), sd = sd(glm1[["residuals"]]))
print(ks_test)

## p-value is > 0.05, so we fail to reject the null hypothesis that the residuals are normally distributed

## The Shapiro-Wilk and Kolmogorov-Smirnov tests produce contradictory results, and the Q-Q plot may indicate
## a mild deviation from normality, but the overall results do not confirm that the residuals severely depart
## from a normal distribution.

```


## MAP
```{r}
library(tmap)
library(tmaptools)
library(leaflet)
library(sf)
library(leaflet.extras)
library(raster)

GAgeo <- st_read("C:\\Users\\bgila\\Downloads\\tl_2020_us_county")
head(GAgeo)

GAmap <- merge(GAgeo, GA2016, by.x = "GEOID", by.y = "GEOID") ## Use GEOID to map each county in Georgia

## initial map test
gastaticmap <- tm_shape(GAmap) +
tm_fill(GAmap$White, title="White", palette = "PRGn") +
tm_borders(alpha=.5) +
tm_style("classic") 

gastaticmap ## WORKED

library(ggplot2)

colnames(GAmap)<-gsub("/", "_", colnames(GAmap)) ## Error had been occurring due to the "/" in certain variables, changing to "_"


lapply(c(19:20, 22:28, 30:34, 36:38, 40:46), function(x) {  ## Select Variables to Map
  
  randint<-runif(1, 0, 3)                         ## helps randomize color in each map
  if (randint >=0 & randint <= 1) {
    low_color<-"#90EE90"
    high_color<-"#006400"
  } else if(randint > 1 & randint <=2) {
    low_color<-"#D9F1FF"
    high_color<-"#009DFF"
  } else {
    low_color<-"#F5D8CE"
    high_color<-"#E18B6D"
  }
  
  df<-GAmap[, c(x, 42)]
  colnames(df)[1]<-"fill_value"
  
  png(paste0("C:\\Users\\bgila\\Desktop\\SPRING2021\\DATA 793\\Project", "\\", colnames(GAmap[, x]), "_map.png")) ## naming PNG files
  
  plot_a<-ggplot(data = df) + 
  geom_sf(aes(fill = fill_value)) +
  scale_fill_gradient(low = low_color, high = high_color)+
  ggtitle(colnames(GAmap[, x]), "by GA county")
  
  print(plot_a)
  
  dev.off()
})

## SHOULD PRINT OUT MAPS FOR ALL 24 VARIABLES INCLUDED

```
